A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-05-31 main.tex
Add six types of judgement.
Add descriptions for every theorem.

@@

In this section, we define the type system of \LMD.
\LMD\ is little complicated because it contains dependent types.
It contains typing, kinding, well-formed kinding, term equality, type equality, and kind equality.

@@

\begin{theorem}[Stage Substituition Lemma]
	\begin{flalign*}
		\text{If\ } \G \V M:\tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]:\tau[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau::K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]::K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\iskind @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon] \iskind @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V M\E N : \tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]\E N[\beta \mapsto \epsilon] : \tau[\beta \mapsto \epsilon]  @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau\E \sigma : K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]\E \sigma[\beta \mapsto \epsilon] : K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\E J @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon]\E J[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&
	\end{flalign*}
\end{theorem}

@@

\begin{theorem}[Inversion Lemma for $\Pi$ type]
	If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$ then
	\begin{enumerate}
		\item $\G \V \sigma \E \sigma'@A$
		\item $\G ,x:\sigma@A\V M:\tau@A$
	\end{enumerate}
	\item If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
	\item If $\G \V (\Pi x:\sigma.\tau) \E \rho : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
\end{theorem}

@@

\begin{theorem}[Preservation]
	If $\G\V M:\tau @A$ and $M \longrightarrow M'$, then $\G\V M':\tau @A$\\
\end{theorem}

@@

\begin{theorem}[Strong Normalization]
	If $\G\V^A M:\tau$ then there is no infinite sequence of terms $(M_i)_{i\ge1}$ and 
	$M_i \longrightarrow M_{i+1}$ for $i\ge 1$
\end{theorem}

@@

\begin{theorem}[Confluence]
	For any term $M$, if $M \longrightarrow^* M'$ and $M \longrightarrow^* M''$ then
	there exists $M'''$ that satisfies $M' \longrightarrow^* M'''$ and $M'' \longrightarrow^* M'''$.
\end{theorem}

@@

\begin{theorem}[Progress]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow M'$.
\end{theorem}

@@

>>>>

\subsection{Dependent Types}

% Dependent Types

% Kind, Well-formed kind

% Type Equality

@@

In this section, we define the type system of \LMD.
\LMD\ is little complicated because it contains dependent types.
It contains typing, kinding, well-formed kinding, term equality, type equality, and kind equality.
In other words, there are six types of judgement in \LMD.
We show them in Figure \ref{fig:LMD-six-judgements}.

\begin{figure}
	\begin{center}
		\begin{align*}
			\G & \V M : \tau @ A \\
			\G & \V \tau :: K @ A \\
			\G & \V K \iskind @ A \\
			\G & \V M \E N @ A \\
			\G & \V \tau \E \sigma \\
			\G & \V K \E J @ A
		\end{align*}
		\caption{Six types of judgement in \LMD}
		\label{fig:LMD-six-judgements}
	\end{center}
\end{figure}

@@

We need to prove another substitution lemma because there are stage variables.
\begin{theorem}[Stage Substituition Lemma]
	\begin{flalign*}
		\text{If\ } \G \V M:\tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]:\tau[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau::K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]::K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\iskind @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon] \iskind @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V M\E N : \tau @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V M[\beta \mapsto \epsilon]\E N[\beta \mapsto \epsilon] : \tau[\beta \mapsto \epsilon]  @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V \tau\E \sigma : K @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V \tau[\beta \mapsto \epsilon]\E \sigma[\beta \mapsto \epsilon] : K[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&\\
		\text{If\ } \G \V K\E J @A
		&\text{\ then\ } \G[\beta \mapsto \epsilon]\V K[\beta \mapsto \epsilon]\E J[\beta \mapsto \epsilon] @A[\beta \mapsto \epsilon].&
	\end{flalign*}
\end{theorem}

@@

Following three inversion lemmas are needed in proving later theorems.
\begin{lemma}[Inversion Lemma for $\Pi$ type]
	If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$ then
	\begin{enumerate}
		\item $\G \V \sigma \E \sigma'@A$
		\item $\G ,x:\sigma@A\V M:\tau@A$
	\end{enumerate}
	\item If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
	\item If $\G \V (\Pi x:\sigma.\tau) \E \rho : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
\end{lemma}

@@

Reductions preserve typing.
\begin{theorem}[Preservation]
	If $\G\V M:\tau @A$ and $M \longrightarrow M'$, then $\G\V M':\tau @A$\\
\end{theorem}

@@

No typed term has an infinite reduction sequence.
\begin{theorem}[Strong Normalization]
	If $\G\V^A M:\tau$ then there is no infinite sequence of terms $(M_i)_{i\ge1}$ and 
	$M_i \longrightarrow M_{i+1}$ for $i\ge 1$
\end{theorem}

@@

Any reduction sequences from one typed term converge.
Because we have proved Strong Normalization, we can use Newman's Lemma to prove Confluence.
\begin{theorem}[Confluence]
	For any term $M$, if $M \longrightarrow^* M'$ and $M \longrightarrow^* M''$ then
	there exists $M'''$ that satisfies $M' \longrightarrow^* M'''$ and $M'' \longrightarrow^* M'''$.
\end{theorem}

@@

Every typed term is a value or can be reduced to another typed term.
\begin{theorem}[Progress]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow M'$.
\end{theorem}

@@

For every typed term, we can find just one redex to reduce by the evaluation context or it is a value.
This theorem is important because it ensure 
that the evaluation context decides a redex to reduce deterministically.
Specifically, this theorem guarantee that 
when you write a interpreter using the evaluation context of \LMD,
your interpreter works just as intended.
